---
title: "Initial Data Processing, Exp24 + Exp27 combined"
author: "Jim Sanford"
date: "2026-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MSnSet.utils)
library(tidyverse)
library(eulerr)
library(ggpubr)
library(readr)
```

## Data

We have reprocessed the Exp 24 and Exp 27 data together (rather than separate) thru DIA-NN or Spectronaut. There are multiple versions of the data (settings in DIA-NN related to FAIMS CV correction), and we need to evaluate these versions side-by-side to identify the best set to move forward with. This first notebook will clean the data and write out tables to be used in subsequence notebooks.

```{r import and clean data, message=FALSE}
##Import protein-level data taken from DIA-NN pg matrix output. see log file for parameter information. 
##Processing according to Sam/Camilo's previous work in Exp24 and Exp27
##Files generated through PNNL DMS DIA-NN analysis and James Fulcher December 2025

#Standard DMS DIA-NN workflow; also stored as syn72325476
DIA_combo_v1 <-read_tsv("data/Old_Approach_Med_Normalized.tsv")

#New DIA-NN FAIMS workflow from James; also stored as syn72325477
DIA_combo_v2 <-read_tsv("data/New_Approach_Med_Normalized.tsv")

#New DIAn-NN FAIMS workflow with QuantUMS from James; also stored as syn72325478
DIA_combo_v3 <-read_tsv("data/New_Approach_Med_Normalized_QuantumMS.tsv")

#New Spectronaut data from James for comparison, also stored as syn72325479
DIA_combo_v4 <- read_tsv("data/Spectronaut_MedNorm.tsv")

##remove contaminants
DIA_combo_v1 <-DIA_combo_v1 %>% filter(!str_detect(Protein.Group,"Contaminant|contaminant"))
DIA_combo_v2 <-DIA_combo_v2 %>% filter(!str_detect(Protein.Group,"Contaminant|contaminant"))
DIA_combo_v3 <-DIA_combo_v3 %>% filter(!str_detect(Protein.Group,"Contaminant|contaminant"))
DIA_combo_v4 <-DIA_combo_v4 %>% filter(!str_detect(Protein.Group,"Contaminant|contaminant"))

##change Protein.Group to row names
DIA_combo_v1 <- DIA_combo_v1 %>% column_to_rownames("Protein.Group")
DIA_combo_v2 <- DIA_combo_v2 %>% column_to_rownames("Protein.Group")
DIA_combo_v3 <- DIA_combo_v3 %>% column_to_rownames("Protein.Group")
DIA_combo_v4 <- DIA_combo_v4 %>% column_to_rownames("Protein.Group")

## Save crosstabs
write.table(DIA_combo_v1, file = "crosstab_PTRC_combo_old_approach.txt",
            sep = "\t", quote = FALSE, row.names = TRUE,col.names = TRUE)

write.table(DIA_combo_v2, file = "crosstab_PTRC_combo_new_approach.txt",
            sep = "\t", quote = FALSE, row.names = TRUE,col.names = TRUE)

write.table(DIA_combo_v3, file = "crosstab_PTRC_combo_new_approach_quantumMS.txt",
            sep = "\t", quote = FALSE, row.names = TRUE,col.names = TRUE)

write.table(DIA_combo_v4, file = "crosstab_PTRC_combo_SpectroNaut.txt",
            sep = "\t", quote = FALSE, row.names = TRUE,col.names = TRUE)
```

```{r Venn diagrams, message=FALSE}
##If desired for visualization of protein ID overlaps, Venn diagrams can be made from the rownames of any given datasets
p <- plot(euler(list("Original"=rownames(DIA_combo_v1),
                     "Updated"=rownames(DIA_combo_v3))), 
          fills=list(fill=c('#0072b2','#d55e00'), 
                     alpha=c(0.5,0.5)),quantities = list(type = "counts"),
          main="Protein ID Overlap: Exp 24 +27 Analysis Approaches",
          eulerr_options(labels = list(col = "black", fontsize = 20),
                         quantities = list(col = "black", fontsize=20),
                         main = list(col = "black", fontsize=16)))
print(p)
```

```{r generate files for protein-to-gene mapping, message=FALSE}
##Tally how many rows aren't unique mappings (multiple proteins separated by ":")
crosstab_prot_v1 <- DIA_combo_v1
table(grepl(";", rownames(crosstab_prot_v1)))

crosstab_prot_v2 <- DIA_combo_v2
table(grepl(";", rownames(crosstab_prot_v2)))

crosstab_prot_v3 <- DIA_combo_v3
table(grepl(";", rownames(crosstab_prot_v3)))

crosstab_prot_v4 <- DIA_combo_v4
table(grepl(";", rownames(crosstab_prot_v4)))

##Remove these non-unique rows
crosstab_prot_v1 <- crosstab_prot_v1[!grepl(";", rownames(crosstab_prot_v1)), ]
crosstab_prot_v2 <- crosstab_prot_v2[!grepl(";", rownames(crosstab_prot_v2)), ]
crosstab_prot_v3 <- crosstab_prot_v3[!grepl(";", rownames(crosstab_prot_v3)), ]
crosstab_prot_v4 <- crosstab_prot_v4[!grepl(";", rownames(crosstab_prot_v4)), ]

##Write out the row names for mapping via UniProt
write.table(rownames(crosstab_prot_v1), quote = F, file = "dia_sorted_cells_proteins_v1.txt", col.names = F, row.names = F)
write.table(rownames(crosstab_prot_v2), quote = F, file = "dia_sorted_cells_proteins_v2.txt", col.names = F, row.names = F)
write.table(rownames(crosstab_prot_v3), quote = F, file = "dia_sorted_cells_proteins_v3.txt", col.names = F, row.names = F)
write.table(rownames(crosstab_prot_v4), quote = F, file = "dia_sorted_cells_proteins_v4.txt", col.names = F, row.names = F)

###Go here to map these to gene names: https://www.uniprot.org/id-mapping 
###(uniprotKB AC/ID to Gene Name (select under uniprot))
```

```{r map proteins to genes and collapse, message=FALSE}
##Import the mapping tables downloaded from UniProt 
##Will upload these to synapse and update code rather than having to use local files
conv_df_v1 <- read.table("data/idmapping_2025_12_19_v1.tsv", sep = "\t", header = T) %>%
  group_by(From) %>%
  mutate(total = n()) %>% ungroup() %>%
  group_by(To) %>% mutate(total_2 = n()) %>% ungroup()

conv_df_v1 <- conv_df_v1 %>%
  group_by(From) %>% slice(1)

conv_df_v2 <- read.table("data/idmapping_2025_12_19_v2.tsv", sep = "\t", header = T) %>%
  group_by(From) %>%
  mutate(total = n()) %>% ungroup() %>%
  group_by(To) %>% mutate(total_2 = n()) %>% ungroup()

conv_df_v2 <- conv_df_v2 %>%
  group_by(From) %>% slice(1)

conv_df_v3 <- read.table("data/idmapping_2025_12_19_v3.tsv", sep = "\t", header = T) %>%
  group_by(From) %>%
  mutate(total = n()) %>% ungroup() %>%
  group_by(To) %>% mutate(total_2 = n()) %>% ungroup()

conv_df_v3 <- conv_df_v3 %>%
  group_by(From) %>% slice(1)

conv_df_v4 <- read.table("data/idmapping_2026_01_05.tsv", sep = "\t", header = T) %>%
  group_by(From) %>%
  mutate(total = n()) %>% ungroup() %>%
  group_by(To) %>% mutate(total_2 = n()) %>% ungroup()

conv_df_v4 <- conv_df_v4 %>%
  group_by(From) %>% slice(1)

###Note: do these steps for each of the datasets by changing file names

##Merge these mappings with prior crosstab to exclude anything that doesn't map to gene name
crosstab_df_v1 <- crosstab_prot_v1 %>% as.data.frame() %>%
  mutate(From = rownames(crosstab_prot_v1)) %>%
  merge(conv_df_v1, by = "From") %>%
  mutate(feature = To) %>% select(-From, -To, -total, -total_2)

##log transform data then take the mean samplewise if two proteins map to the same gene
##Make sure to include na.rm=TRUE in the mean function!!!
crosstab_df_v1 <- crosstab_df_v1 %>%
  tidyr::pivot_longer(-feature, values_to = "value", names_to = "sample") %>%
  mutate(value = log2(value)) %>%
  tidyr::pivot_wider(names_from = "sample", values_from = "value", values_fn = ~mean(.x, na.rm=TRUE))

##Write tables if desired (log2-transformed data, no missingness filters)
write.table(crosstab_df_v1, file = "sorted_cells_dia_5469_6599_combo_v1_unfiltered.txt", sep = "\t", quote = F, row.names = F)
```

```{r evaluate/filter for missingness and median-center data, message=FALSE}
###Note: as above, these steps need to be performed dataset-by-dataset by changing file names (v1, v2, v3, v4)

#Make data a matrix with features as rownames
crosstab_v1 <- crosstab_df_v1 %>%
  column_to_rownames("feature") %>% 
  as.matrix()

#Plot missingness and data distribution
hist(rowSums(is.na(crosstab_v1)))

boxplot(crosstab_v1)

#Calculate column medians and median-center data
col_medians <- apply(crosstab_v1, 2, median, na.rm = T)
crosstab_v1_mc <- sweep(crosstab_v1, 2, col_medians, FUN = '-')

boxplot(crosstab_v1_mc)

#Write table: median-centered, all proteins (no missingness filter)
write.table(crosstab_v1_mc, "sorted_cells_dia_5469_6599_combo_v1_allquant.txt", sep = "\t", quote = F)

#Now make a version where proteins are filtered for <50% missingness prior to median-centering
#Do this by building an msnset but there's probably an easier way
meta_v1 <- data.frame(sample = colnames(crosstab_v1)) %>%
  mutate(group = sub("^.*_", "", sample),
         patient = sub("_.*$", "", sample))
rownames(meta_v1) <- meta_v1$sample

m1 <- MSnSet(exprs = crosstab_v1, pData = meta_v1)

#Filter for proteins with <50% missingness, plot missingness and distribution
m1 <- m1[rowSums(is.na(exprs(m1))) < (ncol(m1)/2), ]

hist(rowSums(is.na(exprs(m1))))

boxplot(exprs(m1))

#Calculate column medians and median-center data
col_medians <- apply(exprs(m1), 2, median, na.rm = T)
exprs(m1) <- sweep(exprs(m1), 2, col_medians, FUN = '-')

boxplot(exprs(m1))

#Write table: median-centered, only proteins with <50% missingness
write.table(exprs(m1), "sorted_cells_dia_5469_6599_combo_v1_minhalfsampquant.txt", sep = "\t", quote = F)

```

```{r last evaluations: missingness within samples, %CV of different datasets message=FALSE}
##Check missingness within samples (to determine if some samples should be removed later)
x1 <- as.data.frame(colMeans(!is.na(exprs(m1))))
x2 <- as.data.frame(colMeans(!is.na(exprs(m2))))
x3 <- as.data.frame(colMeans(!is.na(exprs(m3))))
x4 <- as.data.frame(colMeans(!is.na(exprs(m4))))

##%CV calculations (after data has been log2-transformed and mapped to/collapsed at the gene level)
calculate_cv <- function(x) {
  # Remove any NA values to avoid calculation errors
  x <- na.omit(x)
  
  # Compute mean and standard deviation
  mean_value <- mean(x)
  sd_value <- sd(x)
  
  # Handle edge case: Avoid division by zero if mean is zero
  if (mean_value == 0) {
    return(NA) # Return NA for undefined %CV
  }
  
  # Calculate the %CV
  cv <- (sd_value / mean_value) * 100
  return(cv)
}

cvs_v1 <- apply(crosstab_v1, 1, calculate_cv)
cvs_v2 <- apply(crosstab_v2, 1, calculate_cv)
cvs_v3 <- apply(crosstab_v3, 1, calculate_cv)
cvs_v4 <- apply(crosstab_v4, 1, calculate_cv)


cv_list <- list(v1 = cvs_v1,
                v2 = cvs_v2,
                v3 = cvs_v3,
                v4 = cvs_v4)

boxplot(cv_list, main = "%CV, log2-transformed Data")
```
