---
title: "Initial Data Processing, Exp24 + Exp27 combined"
author: "Jim Sanford"
date: "2026-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MSnSet.utils)
library(tidyverse)
library(eulerr)
library(ggpubr)
library(readr)
library(synapser)
```

## Data Processing: DIA-NN Results tables to crosstabs for downstream analysis

We have reprocessed the Exp 24 and Exp 27 data together (rather than separate) thru DIA-NN or Spectronaut. There are multiple versions of the data (settings in DIA-NN related to FAIMS CV correction), and we need to evaluate these versions side-by-side to identify the best set to move forward with. This first notebook will clean the data and write out tables to be used in subsequent notebooks.

This first section reads in DIA-NN results tables (stored on Synapse), removes contaminants, and writes out crosstabs.

```{r import and clean data, message=FALSE}
##Import protein-level data taken from DIA-NN pg matrix output. see log file for parameter information. 
##Processing according to Sam/Camilo's previous work in Exp24 and Exp27
##Files generated through PNNL DMS DIA-NN analysis and James Fulcher December 2025

#For pulling files from synapse
synapser::synLogin()

#Standard DMS DIA-NN workflow;  stored as syn72325476
v1 <- synapser::synGet("syn72325476")
DIA_combo_v1 <- read_tsv(v1$path)

#New DIA-NN FAIMS workflow from James;  stored as syn72325477
v2 <- synapser::synGet("syn72325477")
DIA_combo_v2 <- read_tsv(v2$path)

#New DIAn-NN FAIMS workflow with QuantUMS from James;  stored as syn72325478
v3 <- synapser::synGet("syn72325478")
DIA_combo_v3 <- read_tsv(v3$path)

#New Spectronaut data from James for comparison, stored as syn72325479
v4 <- synapser::synGet("syn72325479")
DIA_combo_v4 <- read_tsv(v4$path)

##remove contaminants
DIA_combo_v1 <-DIA_combo_v1 %>% filter(!str_detect(Protein.Group,"Contaminant|contaminant"))
DIA_combo_v2 <-DIA_combo_v2 %>% filter(!str_detect(Protein.Group,"Contaminant|contaminant"))
DIA_combo_v3 <-DIA_combo_v3 %>% filter(!str_detect(Protein.Group,"Contaminant|contaminant"))
DIA_combo_v4 <-DIA_combo_v4 %>% filter(!str_detect(Protein.Group,"Contaminant|contaminant"))

##change Protein.Group to row names
DIA_combo_v1 <- DIA_combo_v1 %>% column_to_rownames("Protein.Group")
DIA_combo_v2 <- DIA_combo_v2 %>% column_to_rownames("Protein.Group")
DIA_combo_v3 <- DIA_combo_v3 %>% column_to_rownames("Protein.Group")
DIA_combo_v4 <- DIA_combo_v4 %>% column_to_rownames("Protein.Group")

## Save crosstabs
write.table(DIA_combo_v1, file = "crosstab_PTRC_combo_old_approach.txt",
            sep = "\t", quote = FALSE, row.names = TRUE,col.names = TRUE)

write.table(DIA_combo_v2, file = "crosstab_PTRC_combo_new_approach.txt",
            sep = "\t", quote = FALSE, row.names = TRUE,col.names = TRUE)

write.table(DIA_combo_v3, file = "crosstab_PTRC_combo_new_approach_quantumMS.txt",
            sep = "\t", quote = FALSE, row.names = TRUE,col.names = TRUE)

write.table(DIA_combo_v4, file = "crosstab_PTRC_combo_SpectroNaut.txt",
            sep = "\t", quote = FALSE, row.names = TRUE,col.names = TRUE)
```

This section allows for generation of Venn diagrams to look at overlap of protein group IDs across any of the datasets desired.

```{r Venn diagrams, message=FALSE}
##If desired for visualization of protein ID overlaps, Venn diagrams can be made from the rownames of any given datasets
p <- plot(euler(list("Original"=rownames(DIA_combo_v1),
                     "Updated"=rownames(DIA_combo_v3))), 
          fills=list(fill=c('#0072b2','#d55e00'), 
                     alpha=c(0.5,0.5)),quantities = list(type = "counts"),
          main="Protein ID Overlap: Exp 24 +27 Analysis Approaches",
          eulerr_options(labels = list(col = "black", fontsize = 20),
                         quantities = list(col = "black", fontsize=20),
                         main = list(col = "black", fontsize=16)))
print(p)
```

Now, non-unique rows are removed from the crosstab, and the protein IDs (UniProt accessions from DIA-NN search) from each dataset are written out as tables. These tables are then mapped to gene symbols via the UniProt website.

```{r generate files for protein-to-gene mapping, message=FALSE}
##Tally how many rows aren't unique mappings (multiple proteins separated by ":")
crosstab_prot_v1 <- DIA_combo_v1
table(grepl(";", rownames(crosstab_prot_v1)))

crosstab_prot_v2 <- DIA_combo_v2
table(grepl(";", rownames(crosstab_prot_v2)))

crosstab_prot_v3 <- DIA_combo_v3
table(grepl(";", rownames(crosstab_prot_v3)))

crosstab_prot_v4 <- DIA_combo_v4
table(grepl(";", rownames(crosstab_prot_v4)))

##Remove these non-unique rows
crosstab_prot_v1 <- crosstab_prot_v1[!grepl(";", rownames(crosstab_prot_v1)), ]
crosstab_prot_v2 <- crosstab_prot_v2[!grepl(";", rownames(crosstab_prot_v2)), ]
crosstab_prot_v3 <- crosstab_prot_v3[!grepl(";", rownames(crosstab_prot_v3)), ]
crosstab_prot_v4 <- crosstab_prot_v4[!grepl(";", rownames(crosstab_prot_v4)), ]

##Write out the row names for mapping via UniProt
write.table(rownames(crosstab_prot_v1), quote = F, file = "dia_sorted_cells_proteins_v1.txt", col.names = F, row.names = F)
write.table(rownames(crosstab_prot_v2), quote = F, file = "dia_sorted_cells_proteins_v2.txt", col.names = F, row.names = F)
write.table(rownames(crosstab_prot_v3), quote = F, file = "dia_sorted_cells_proteins_v3.txt", col.names = F, row.names = F)
write.table(rownames(crosstab_prot_v4), quote = F, file = "dia_sorted_cells_proteins_v4.txt", col.names = F, row.names = F)

###Go here to map these to gene names: https://www.uniprot.org/id-mapping 
###(uniprotKB AC/ID to Gene Name (select under uniprot))
```

Protein-level crosstabs are now mapped to gene symbols using the protein-to-gene mappings from UniProt (now stored on synapse). Proteins that don't map to a gene symbol are excluded. For instances where multiple UniProt accessions map to the same gene symbol, the mean is calculated and retained. Data is then log2-transformed, and these crosstabs can be written out if desired.

```{r map proteins to genes and collapse, message=FALSE}
##Import the mapping tables downloaded from UniProt and stored on synapse
ids_v1 <- synapser::synGet("syn74022069")
conv_df_v1 <- read.table(ids_v1$path, sep = "\t", header = T) %>%
  group_by(From) %>%
  mutate(total = n()) %>% ungroup() %>%
  group_by(To) %>% mutate(total_2 = n()) %>% ungroup()

conv_df_v1 <- conv_df_v1 %>%
  group_by(From) %>% slice(1)

ids_v2 <- synapser::synGet("syn74022072")
conv_df_v2 <- read.table(ids_v2$path, sep = "\t", header = T) %>%
  group_by(From) %>%
  mutate(total = n()) %>% ungroup() %>%
  group_by(To) %>% mutate(total_2 = n()) %>% ungroup()

conv_df_v2 <- conv_df_v2 %>%
  group_by(From) %>% slice(1)

ids_v3 <- synapser::synGet("syn74022073")
conv_df_v3 <- read.table(ids_v3$path, sep = "\t", header = T) %>%
  group_by(From) %>%
  mutate(total = n()) %>% ungroup() %>%
  group_by(To) %>% mutate(total_2 = n()) %>% ungroup()

conv_df_v3 <- conv_df_v3 %>%
  group_by(From) %>% slice(1)

ids_v4 <- synapser::synGet("syn74022080")
conv_df_v4 <- read.table(ids_v4$path, sep = "\t", header = T) %>%
  group_by(From) %>%
  mutate(total = n()) %>% ungroup() %>%
  group_by(To) %>% mutate(total_2 = n()) %>% ungroup()

conv_df_v4 <- conv_df_v4 %>%
  group_by(From) %>% slice(1)

##Merge these mappings with prior crosstab to exclude anything that doesn't map to gene name
crosstab_df_v1 <- crosstab_prot_v1 %>% as.data.frame() %>%
  mutate(From = rownames(crosstab_prot_v1)) %>%
  merge(conv_df_v1, by = "From") %>%
  mutate(feature = To) %>% select(-From, -To, -total, -total_2)

crosstab_df_v2 <- crosstab_prot_v2 %>% as.data.frame() %>%
  mutate(From = rownames(crosstab_prot_v2)) %>%
  merge(conv_df_v2, by = "From") %>%
  mutate(feature = To) %>% select(-From, -To, -total, -total_2)

crosstab_df_v3 <- crosstab_prot_v3 %>% as.data.frame() %>%
  mutate(From = rownames(crosstab_prot_v3)) %>%
  merge(conv_df_v3, by = "From") %>%
  mutate(feature = To) %>% select(-From, -To, -total, -total_2)

crosstab_df_v4 <- crosstab_prot_v4 %>% as.data.frame() %>%
  mutate(From = rownames(crosstab_prot_v4)) %>%
  merge(conv_df_v4, by = "From") %>%
  mutate(feature = To) %>% select(-From, -To, -total, -total_2)

##log transform data then take the mean samplewise if two proteins map to the same gene
##Make sure to include na.rm=TRUE in the mean function!!!
crosstab_df_v1 <- crosstab_df_v1 %>%
  tidyr::pivot_longer(-feature, values_to = "value", names_to = "sample") %>%
  mutate(value = log2(value)) %>%
  tidyr::pivot_wider(names_from = "sample", values_from = "value", values_fn = ~mean(.x, na.rm=TRUE))

crosstab_df_v2 <- crosstab_df_v2 %>%
  tidyr::pivot_longer(-feature, values_to = "value", names_to = "sample") %>%
  mutate(value = log2(value)) %>%
  tidyr::pivot_wider(names_from = "sample", values_from = "value", values_fn = ~mean(.x, na.rm=TRUE))

crosstab_df_v3 <- crosstab_df_v3 %>%
  tidyr::pivot_longer(-feature, values_to = "value", names_to = "sample") %>%
  mutate(value = log2(value)) %>%
  tidyr::pivot_wider(names_from = "sample", values_from = "value", values_fn = ~mean(.x, na.rm=TRUE))

crosstab_df_v4 <- crosstab_df_v4 %>%
  tidyr::pivot_longer(-feature, values_to = "value", names_to = "sample") %>%
  mutate(value = log2(value)) %>%
  tidyr::pivot_wider(names_from = "sample", values_from = "value", values_fn = ~mean(.x, na.rm=TRUE))

##Write tables if desired (log2-transformed data, no missingness filters)
write.table(crosstab_df_v1, file = "sorted_cells_dia_5469_6599_combo_v1_unfiltered.txt", sep = "\t", quote = F, row.names = F)
write.table(crosstab_df_v2, file = "sorted_cells_dia_5469_6599_combo_v2_unfiltered.txt", sep = "\t", quote = F, row.names = F)
write.table(crosstab_df_v3, file = "sorted_cells_dia_5469_6599_combo_v3_unfiltered.txt", sep = "\t", quote = F, row.names = F)
write.table(crosstab_df_v4, file = "sorted_cells_dia_5469_6599_combo_v4_unfiltered.txt", sep = "\t", quote = F, row.names = F)
```

This section median-centers the data and writes out crosstabs for each dataset. This is done first on the entire dataset, and then a second time on a version of the datasets where proteins with >50% missing values have been removed. Histograms of missingness, and boxplots within each sample before and after median-centering are generated for data visualization.

```{r evaluate/filter for missingness and median-center data, message=FALSE}
#Make data a matrix with features as rownames
crosstab_v1 <- crosstab_df_v1 %>%
  column_to_rownames("feature") %>% 
  as.matrix()

crosstab_v2 <- crosstab_df_v2 %>%
  column_to_rownames("feature") %>% 
  as.matrix()

crosstab_v3 <- crosstab_df_v3 %>%
  column_to_rownames("feature") %>% 
  as.matrix()

crosstab_v4 <- crosstab_df_v4 %>%
  column_to_rownames("feature") %>% 
  as.matrix()

#Plot missingness and data distribution
hist(rowSums(is.na(crosstab_v1)))
boxplot(crosstab_v1)

hist(rowSums(is.na(crosstab_v2)))
boxplot(crosstab_v2)

hist(rowSums(is.na(crosstab_v3)))
boxplot(crosstab_v3)

hist(rowSums(is.na(crosstab_v4)))
boxplot(crosstab_v4)

#Calculate column medians and median-center data
col_medians_v1 <- apply(crosstab_v1, 2, median, na.rm = T)
crosstab_v1_mc <- sweep(crosstab_v1, 2, col_medians_v1, FUN = '-')
boxplot(crosstab_v1_mc)

col_medians_v2 <- apply(crosstab_v2, 2, median, na.rm = T)
crosstab_v2_mc <- sweep(crosstab_v2, 2, col_medians_v2, FUN = '-')
boxplot(crosstab_v2_mc)

col_medians_v3 <- apply(crosstab_v3, 2, median, na.rm = T)
crosstab_v3_mc <- sweep(crosstab_v3, 2, col_medians_v3, FUN = '-')
boxplot(crosstab_v3_mc)

col_medians_v4 <- apply(crosstab_v4, 2, median, na.rm = T)
crosstab_v4_mc <- sweep(crosstab_v4, 2, col_medians_v4, FUN = '-')
boxplot(crosstab_v4_mc)

#Write table: median-centered, all proteins (no missingness filter)
write.table(crosstab_v1_mc, "sorted_cells_dia_5469_6599_combo_v1_allquant.txt", sep = "\t", quote = F)
write.table(crosstab_v2_mc, "sorted_cells_dia_5469_6599_combo_v2_allquant.txt", sep = "\t", quote = F)
write.table(crosstab_v3_mc, "sorted_cells_dia_5469_6599_combo_v3_allquant.txt", sep = "\t", quote = F)
write.table(crosstab_v4_mc, "sorted_cells_dia_5469_6599_combo_v4_allquant.txt", sep = "\t", quote = F)

#Now make a version where proteins are filtered for <50% missingness prior to median-centering
#Do this by building an msnset but there's probably an easier way
meta_v1 <- data.frame(sample = colnames(crosstab_v1)) %>%
  mutate(group = sub("^.*_", "", sample),
         patient = sub("_.*$", "", sample))
rownames(meta_v1) <- meta_v1$sample

m1 <- MSnSet(exprs = crosstab_v1, pData = meta_v1)

meta_v2 <- data.frame(sample = colnames(crosstab_v2)) %>%
  mutate(group = sub("^.*_", "", sample),
         patient = sub("_.*$", "", sample))
rownames(meta_v2) <- meta_v2$sample

m2 <- MSnSet(exprs = crosstab_v2, pData = meta_v2)

meta_v3 <- data.frame(sample = colnames(crosstab_v3)) %>%
  mutate(group = sub("^.*_", "", sample),
         patient = sub("_.*$", "", sample))
rownames(meta_v3) <- meta_v3$sample

m3 <- MSnSet(exprs = crosstab_v3, pData = meta_v3)

meta_v4 <- data.frame(sample = colnames(crosstab_v4)) %>%
  mutate(group = sub("^.*_", "", sample),
         patient = sub("_.*$", "", sample))
rownames(meta_v4) <- meta_v4$sample

m4 <- MSnSet(exprs = crosstab_v4, pData = meta_v4)

#Filter for proteins with <50% missingness, plot missingness and distribution
m1 <- m1[rowSums(is.na(exprs(m1))) < (ncol(m1)/2), ]
hist(rowSums(is.na(exprs(m1))))
boxplot(exprs(m1))

m2 <- m2[rowSums(is.na(exprs(m2))) < (ncol(m2)/2), ]
hist(rowSums(is.na(exprs(m2))))
boxplot(exprs(m2))

m3 <- m3[rowSums(is.na(exprs(m3))) < (ncol(m3)/2), ]
hist(rowSums(is.na(exprs(m3))))
boxplot(exprs(m3))

m4 <- m4[rowSums(is.na(exprs(m4))) < (ncol(m4)/2), ]
hist(rowSums(is.na(exprs(m4))))
boxplot(exprs(m4))

#Calculate column medians and median-center data
col_medians_v1 <- apply(exprs(m1), 2, median, na.rm = T)
exprs(m1) <- sweep(exprs(m1), 2, col_medians_v1, FUN = '-')
boxplot(exprs(m1))

col_medians_v2 <- apply(exprs(m2), 2, median, na.rm = T)
exprs(m2) <- sweep(exprs(m2), 2, col_medians_v2, FUN = '-')
boxplot(exprs(m2))

col_medians_v3 <- apply(exprs(m3), 2, median, na.rm = T)
exprs(m3) <- sweep(exprs(m3), 2, col_medians_v3, FUN = '-')
boxplot(exprs(m3))

col_medians_v4 <- apply(exprs(m4), 2, median, na.rm = T)
exprs(m4) <- sweep(exprs(m4), 2, col_medians_v4, FUN = '-')
boxplot(exprs(m4))

#Write table: median-centered, only proteins with <50% missingness
write.table(exprs(m1), "sorted_cells_dia_5469_6599_combo_v1_minhalfsampquant.txt", sep = "\t", quote = F)
write.table(exprs(m2), "sorted_cells_dia_5469_6599_combo_v2_minhalfsampquant.txt", sep = "\t", quote = F)
write.table(exprs(m3), "sorted_cells_dia_5469_6599_combo_v3_minhalfsampquant.txt", sep = "\t", quote = F)
write.table(exprs(m4), "sorted_cells_dia_5469_6599_combo_v4_minhalfsampquant.txt", sep = "\t", quote = F)
```

This last section looks at data missingness within samples, as previous analysis of this data had removed samples that did not meet a certain coverage threshold (75%). No filtering is done, but these values are calculated for reference. Additionally, protein-level %CV is calculated within each dataset and plotted for comparison.

```{r last evaluations: missingness within samples, %CV of different datasets message=FALSE}
##Check missingness within samples (to determine if some samples should be removed later)
x1 <- as.data.frame(colMeans(!is.na(exprs(m1))))
x2 <- as.data.frame(colMeans(!is.na(exprs(m2))))
x3 <- as.data.frame(colMeans(!is.na(exprs(m3))))
x4 <- as.data.frame(colMeans(!is.na(exprs(m4))))

##%CV calculations (after data has been log2-transformed and mapped to/collapsed at the gene level)
calculate_cv <- function(x) {
  # Remove any NA values to avoid calculation errors
  x <- na.omit(x)
  
  # Compute mean and standard deviation
  mean_value <- mean(x)
  sd_value <- sd(x)
  
  # Handle edge case: Avoid division by zero if mean is zero
  if (mean_value == 0) {
    return(NA) # Return NA for undefined %CV
  }
  
  # Calculate the %CV
  cv <- (sd_value / mean_value) * 100
  return(cv)
}

cvs_v1 <- apply(crosstab_v1, 1, calculate_cv)
cvs_v2 <- apply(crosstab_v2, 1, calculate_cv)
cvs_v3 <- apply(crosstab_v3, 1, calculate_cv)
cvs_v4 <- apply(crosstab_v4, 1, calculate_cv)


cv_list <- list(v1 = cvs_v1,
                v2 = cvs_v2,
                v3 = cvs_v3,
                v4 = cvs_v4)

boxplot(cv_list, main = "%CV, log2-transformed Data")
```
