---
title: "Ex1 Global proteomics"
output: html_document
---

The global proteomics pipeline uses the R package [PlexedPiper](https://github.com/vladpetyuk/PlexedPiper). It also requires a connection to the DMS to access data packages.


```{r, echo=F}
library(knitr)
knitr::opts_chunk$set(message=F, warning=F)
t0 <- Sys.time(); print(t0)
```


```{r setup}
library(PlexedPiper)
library(dplyr)
library(PNNL.DMS.utils)
library(Biostrings)
library(tidyverse)

data_package_num <- 5792
data_folder <- "data/global_data"
crosstab_prefix <- "Chr8_UMN1"

dir.create(data_folder)

```


# Study design

Study design information in PlexedPiper is encoded in three tables: fractions, samples, and references. 
These tables can be made using metadata and should be stored on the DMS before processing.

```{r read_study_design}
fractions <- read.table("study_design/global_study_design/fractions.txt", sep = "\t", header = T)
samples <- read.table("study_design/global_study_design/samples.txt", sep = "\t", header = T)
references <- read.table("study_design/global_study_design/references.txt", sep = "\t", header = T)

```

# Process MS-GF+ data

MS-GF+ data is processed in several steps. First, read MS-GF+ output from the DMS. (This step can take a while).

```{r read_msgf}
msgf_data_path <- file.path(data_folder, "msgfData_original.RData")
if (file.exists(msgf_data_path)) {
  load(msgf_data_path)
} else {
  msnid <- read_msgf_data_from_DMS(data_package_num)
  save(msnid, file=msgf_data_path)
}
show(msnid)

```


## Correct peak selection and peptide FDR
We use the target-decoy search strategy method described in [(Elias 2010)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2922680/). 
Filtering is done first at peptide level, then at protein level, both with max FDR of 1%.

```{r}
## peak selection & Peptide filter
msnid <- MSnID::correct_peak_selection(msnid)   ## optional for phospho
msnid <- PlexedPiper::filter_msgf_data_peptide_level(msnid, 0.01)


```


## Remap accessions

This function remaps UniProt protein accessions to gene symbol.

```{r remap_accessions}
options(timeout = 120)
# get fst_dms_path from:
#debug(path_to_FASTA_used_by_DMS)
path_to_FASTA <- path_to_FASTA_used_by_DMS(data_package_num)

# in debug environment after running line 34: 
# assign("fst_dms_pth", temp_filepath, envir = globalenv())

#fst_dms_pth <- path_to_FASTA_used_by_DMS(data_package_num)
#fst_dms_pth <- "/var/folders/8b/3csd5ptx4_9_qmpg2m3_1h3h0000gn/T//RtmphSdOua/file3045652d6d1f.fasta" # changes every time
#fst_dms_pth <- path_to_FASTA
fst <- readAAStringSet(fst_dms_pth)

conv <- data.frame(NAME = names(fst)) %>%
  mutate(UNIPROT = str_extract(names(fst), "^(\\S)+"),
         SYMBOL = str_extract(names(fst), "GN=(\\S)+")) %>%
  dplyr::select(-NAME) %>%
  mutate(UNIPROT = sub(".*\\|(.*)\\|.*", "\\1", UNIPROT),
         SYMBOL = sub("GN=", "", SYMBOL))

msnid_prot <- msnid

msnid <- remap_accessions_uniprot_to_gene(msnid,
                                          organism_name="Homo sapiens",
                                          conversion_table = conv)

show(msnid)
table(is.na(msnid$accession))
msnid <- apply_filter(msnid, "!is.na(accession)")
show(msnid)

```


## Accession FDR


```{r protein_level_filter }
# before 20240731
#fst_dms_pth_2 <- remap_fasta_entry_names(fst_dms_pth, conv, "^.*\\|(.+?)(-\\d+)?\\|.*") 

# ## peptide density and accession filter
# msnid <- compute_num_peptides_per_1000aa(msnid, fst_dms_pth_2)
# msnid <- filter_msgf_data_protein_level(msnid, 0.01)
# show(msnid)
# 
# 
# msnid_prot <- compute_num_peptides_per_1000aa(msnid_prot, fst_dms_pth)
# msnid_prot <- filter_msgf_data_protein_level(msnid_prot, 0.01)
# show(msnid_prot)

# MSnID object
# Working directory: "."
# #Spectrum Files:  48 
# #PSMs: 296095 at 0.066 % FDR
# #peptides: 54333 at 0.13 % FDR
# #accessions: 6011 at 0.99 % FDR
# MSnID object
# Working directory: "."
# #Spectrum Files:  48 
# #PSMs: 443918 at 0.12 % FDR
# #peptides: 93262 at 0.19 % FDR
# #accessions: 25370 at 0.99 % FDR

# after 20240731
# easier to just do it without the function:
conversion_table <- conv
mySequences <- readAAStringSet(fst_dms_pth)
names(mySequences) <- sub(" .*$", "\\1", names(mySequences))
prot_lengths <- data.frame(seq_name = names(mySequences), 
                           seq_length = width(mySequences), stringsAsFactors = FALSE)
by_ <- "seq_name"
names(by_) <- colnames(conversion_table)[1]
conversion_table <- conversion_table %>% inner_join(prot_lengths, 
                                                    by = by_) %>% group_by_at(colnames(conversion_table)[2]) %>% 
  dplyr::slice(which.max(seq_length))
conversion_table <- conversion_table %>% select(-seq_length) %>% 
  distinct() %>% group_by_at(colnames(conversion_table)[1]) %>% 
  top_n(1, dplyr::desc(!!as.name(colnames(conversion_table)[2])))
conversion_vec <- conversion_table %>% pull(2)
names(conversion_vec) <- conversion_table %>% pull(1)
mySequences <- mySequences[names(conversion_vec)]
names(mySequences) <- conversion_vec[names(mySequences)]
path_to_FASTA_remapped <- gsub("\\.fasta.*", "_remapped.fasta", 
                               fst_dms_pth)
writeXStringSet(mySequences, path_to_FASTA_remapped, compress = grepl("[.]gz$", fst_dms_pth))

## peptide density and accession filter
msnid <- compute_num_peptides_per_1000aa(msnid, path_to_FASTA_remapped)
msnid <- filter_msgf_data_protein_level(msnid, 0.01)
show(msnid)


msnid_prot <- compute_num_peptides_per_1000aa(msnid_prot, fst_dms_pth)
msnid_prot <- filter_msgf_data_protein_level(msnid_prot, 0.01)
show(msnid_prot)


# MSnID object
# Working directory: "."
# #Spectrum Files:  48 
# #PSMs: 393851 at 0.00076 % FDR
# #peptides: 84784 at 0.0024 % FDR
# #accessions: 10245 at 0.02 % FDR
# MSnID object
# Working directory: "."
# #Spectrum Files:  48 
# #PSMs: 443918 at 0.12 % FDR
# #peptides: 93262 at 0.19 % FDR
# #accessions: 25370 at 0.99 % FDR

```


## Parsimonious inference

To reduce number of protein identifications, we use a parsimonious inference algorithm described in 
[(Zhang et al. 2007)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2810678/).

```{r parsimonious_inference}
msnid <- infer_parsimonious_accessions(msnid, unique_only=FALSE)
msnid_prot <- infer_parsimonious_accessions(msnid_prot, unique_only=FALSE)
show(msnid)
show(msnid_prot)

# before 20240731
# MSnID object
# Working directory: "."
# #Spectrum Files:  48 
# #PSMs: 290470 at 0.064 % FDR
# #peptides: 53508 at 0.11 % FDR
# #accessions: 5536 at 0.76 % FDR
# MSnID object
# Working directory: "."
# #Spectrum Files:  48 
# #PSMs: 425319 at 0.12 % FDR
# #peptides: 89694 at 0.18 % FDR
# #accessions: 10180 at 1.3 % FDR

# after 20240731
# MSnID object
# Working directory: "."
# #Spectrum Files:  48 
# #PSMs: 386857 at 0.00078 % FDR
# #peptides: 83683 at 0.0024 % FDR
# #accessions: 9162 at 0.022 % FDR
# MSnID object
# Working directory: "."
# #Spectrum Files:  48 
# #PSMs: 425319 at 0.12 % FDR
# #peptides: 89694 at 0.18 % FDR
# #accessions: 10180 at 1.3 % FDR

```


## Remove Decoys

```{r}
msnid <- apply_filter(msnid, "!isDecoy")
show(msnid)

msnid_prot <- apply_filter(msnid_prot, "!isDecoy")
show(msnid_prot)

save(msnid, file=file.path(data_folder, "msgfData_filtered.RData"))

# before 20240731
# MSnID object
# Working directory: "."
# #Spectrum Files:  48 
# #PSMs: 290285 at 0 % FDR
# #peptides: 53447 at 0 % FDR
# #accessions: 5494 at 0 % FDR
# MSnID object
# Working directory: "."
# #Spectrum Files:  48 
# #PSMs: 424818 at 0 % FDR
# #peptides: 89531 at 0 % FDR
# #accessions: 10046 at 0 % FDR

# after 20240731
# MSnID object
# Working directory: "."
# #Spectrum Files:  48 
# #PSMs: 386854 at 0 % FDR
# #peptides: 83681 at 0 % FDR
# #accessions: 9160 at 0 % FDR
# MSnID object
# Working directory: "."
# #Spectrum Files:  48 
# #PSMs: 424818 at 0 % FDR
# #peptides: 89531 at 0 % FDR
# #accessions: 10046 at 0 % FDR

```


## Process MASIC data

Output from the MASIC software is read from DMS, then filtered by interference score.


```{r read_masic}
masic_data_path <- file.path(data_folder, "masicData_original.RData")

options(timeout = 120)
if (file.exists(masic_data_path)) {
  load(masic_data_path)
} else {
  masic_data <- read_masic_data_from_DMS(data_package_num,
                                       interference_score = TRUE)
  save(masic_data, file=masic_data_path)
}

nrow(masic_data)
```


```{r filter_masic}
masic_data <- filter_masic_data(masic_data,
                                interference_score_threshold = 0.5,
                                s2n_threshold = 0)
save(masic_data, file=file.path(data_folder, "masicData_filtered.RData"))

nrow(masic_data)
```


## Create crosstab

The quantitative crosstab combines MS/MS identifications with reporter ion intensities. 
Abundances are taken relative to the reference channel and then log-transformed.


```{r create_global_gene_crosstab}
aggregation_level <- c("accession")

# remove non-samples (empty), but keep references
samples <- samples[samples$ReporterAlias != "EMPTY", ]

debug(create_crosstab)
# once in debug environment, run:
# debug(converting_to_relative_to_reference)
# because there is more than 1 reference on the plex, after line 35 run: 
# quant_data_i_w <- dcast(quant_data_i, Specie ~ ReporterAlias, fun.aggregate = mean, value.var = "Abundance")

quant_cross_tab <- create_crosstab(msnid, 
                                   masic_data, 
                                   aggregation_level, 
                                   fractions, samples, references)

dim(quant_cross_tab)

write.table(quant_cross_tab, file=file.path(data_folder, paste(crosstab_prefix, "crosstab_global_gene_original.txt", sep="_")),
            quote=F, sep="\t")
# before 20240731:
# 5384   66

# after 20240731:
# 8736   66
```




```{r create_global_peptide_crosstab}
aggregation_level <- c("peptide")
quant_cross_tab <- create_crosstab(msnid, 
                                   masic_data, 
                                   aggregation_level, 
                                   fractions, samples, references)
dim(quant_cross_tab)

xx <- psms(msnid) %>% select(accession, peptide, pepSeq) %>% unique()
crosstab <- quant_cross_tab %>% as.data.frame()
crosstab$peptide <- rownames(crosstab)

crosstab <- left_join(crosstab, xx, by = "peptide")
write.table(crosstab %>% select(peptide, pepSeq, accession, everything()), 
            file=file.path(data_folder, paste(crosstab_prefix, "crosstab_global_peptide_original.txt", sep="_")),
            quote=F, sep="\t")
```


```{r create_global_gene_crosstab}
psms(msnid_prot)$accession <- sub("^.*\\|(.*)\\|.*$", "\\1", psms(msnid_prot)$accession)

psms(msnid_prot) <- psms(msnid_prot) %>%
  filter(!grepl("Contaminant", accession))

aggregation_level <- c("accession")
quant_cross_tab_prot <- create_crosstab(msnid_prot, 
                                        masic_data, 
                                        aggregation_level, 
                                        fractions, samples, references)
dim(quant_cross_tab_prot)

write.table(quant_cross_tab_prot, file=file.path(data_folder, paste(crosstab_prefix, "crosstab_global_prot_original.txt", sep="_")),
            quote=F, sep="\t")

# before 20240731:
# 8736   66

# after 20240731:
# 9633   66
```


