---
title: "Cell type markers"
author: "Sara Gosline"
date: "2026-05-21"
output: html_document
---


Here we look at the expression of known markers in the NRAS ASO dataset

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(synapser)
library(ggplot2)
```

## Pull data
First we pull the data that has been put onto synapse.

```{r cars}
synLogin()

global.inputs <- list("Patient ASO" = "syn66726954",
                      "Cell Line ASO" = "syn66726844",
                      "Cell Line KO" = "syn66727085",
                      "Cell Line ASO 2HR" = "syn66727770",
                      "Cell Line ASO 6HR" = "syn66729260",
                      "Cell Line ASO 24HR" = "syn66730823")

glob <- do.call(rbind,lapply(names(global.inputs), function(s){
  tab<-readr::read_csv(synapser::synGet(global.inputs[[s]])$path)|>
    mutate(cond=s)
  return(tab)
}))

glob$signif <-ifelse(glob$adj.P.Val<0.05, "*","")

phospho.inputs <- list("Patient ASO" = "syn66726980",
                       "Cell Line ASO" = "syn66726866",
                       "Cell Line KO" = "syn66727197",
                       "Cell Line ASO 2HR" = "syn66727869",
                       "Cell Line ASO 6HR" = "syn66729468",
                       "Cell Line ASO 24HR" = "syn66730918")

phos <- do.call(rbind,lapply(names(phospho.inputs), function(s){
  tab<-readr::read_csv(synapser::synGet(phospho.inputs[[s]])$path)|>
    mutate(cond=s)
  return(tab)
}))

phos$signif <-ifelse(phos$adj.P.Val<0.05, "*","")

#all.inputs <- list("Global" = global.inputs, "Phospho" = phospho.inputs)

```

## Identify proteins

Elie asked for a few sets of proteins:

monocytic markers: CD11b, CD 14, CD16, CD64, CD68, CD56
primitive markers: CD34, CD117
Other potentially interesting markers: CD33, CD38, CD123

```{r markers, echo=FALSE}

mono <- c('CD11B','CD14','CD16','CD64','CD58','CD46','ITGAM','FCGR3A','FCGR3B','FCGR1A','NCAM1','HLADRA','HLADRB')
prim <- c('CD34','CD117','KIT')
int <- c('CD33','CD38','CD123','IL3RA')

monDiff <- subset(glob, Gene%in%mono)
print(nrow(monDiff))
primDiff <- subset(glob,Gene%in%prim)
print(nrow(primDiff))
intDiff <- subset(glob, Gene%in%int)
print(nrow(intDiff))
```
So there are no primitive markers in the data, they are all monocytic

```{r FC values}

p1<-ggplot(monDiff,aes(x=Gene, y = Log2FC, fill=cond)) +geom_bar(stat='identity',position='dodge') +ggrepel::geom_text_repel(aes(label=signif))

p2<-ggplot(intDiff,aes(x=Gene, y = Log2FC, fill=cond)) +geom_bar(stat='identity',position='dodge') +ggrepel::geom_text_repel(aes(label=signif))

cowplot::plot_grid(p1,p2,nrow=2)
ggsave("cellMarkerPlots.pdf")
readr::write_csv(rbind(monDiff,intDiff),'markerLogFC.csv')

```

## Phospho data

Can we find the proteins in the phospho data?

```{r phospho markers}
phos_mono <- phos[unlist(lapply(mono, function(x) grep(paste(x,'-',sep=''),phos$SUB_SITE))),]
print(dim(phos_mono))

phos_prim <-phos[unlist(lapply(prim, function(x) grep(paste(x,'-',sep=''),phos$SUB_SITE))),]

print(dim(phos_prim))

phos_int <- phos[unlist(lapply(int, function(x) grep(paste(x,'-',sep=''),phos$SUB_SITE))),]
print(dim(phos_int))

```

Now we can plot the phospho data

```{r phospho plots}


p1 <- ggplot(phos_mono,aes(x=SUB_SITE, y = Log2FC, fill=cond)) +geom_bar(stat='identity',position='dodge') +
  ggrepel::geom_text_repel(aes(label=signif)) +
  coord_flip()

p2 <- ggplot(phos_int,aes(x=SUB_SITE, y = Log2FC, fill=cond)) +geom_bar(stat='identity',position='dodge') +
  ggrepel::geom_text_repel(aes(label=signif)) + 
  coord_flip()

p3 <- ggplot(phos_prim,aes(x=SUB_SITE, y = Log2FC, fill=cond)) +geom_bar(stat='identity',position='dodge') +
  ggrepel::geom_text_repel(aes(label=signif)) +
  coord_flip()

cowplot::plot_grid(plotlist = c(p1,p2,p3),nrow=3)

ggsave("phos_cellMarkerPlots.pdf")
readr::write_csv(rbind(phos_mono,phos_int,phos_prim),'phospho_markerLogFC.csv')


```
## Now can we plot raw abundances?

This requires going into the three subdirectories (`exp20`,`exp21`,and `exp22`) and processing the batch correction to save the files locally.
THen we can get all the data into a single table

```{r get raw data}

msns_to_long <- function(eset, dtype = 'global'){
  md <- pData(eset) |>
    tibble::rownames_to_column('tube')
  pd <- exprs(eset) |>
     as.data.frame() |>
     tibble::rownames_to_column('feature') |>
     tidyr::pivot_longer(cols = c(-feature),
                         names_to = 'tube',
                         values_to = 'expr')
  pd |>
    left_join(md) |>
    mutate(data_type = dtype)
}

#exp20
e20g <- readRDS('exp20/exp_20_global.rds') |>
  msns_to_long()
e20p <- readRDS("exp20/exp_20_phospho.rds") |>
  msns_to_long(dtype='phospho')

e20 <- rbind(e20g,e20p)
e20$Treatment = e20$bgd_
e20$Treatment[e20$Treatment=='baseline'] <- 'Control'
e20$experiment <- 'Gilteritinib resistant cell line'

#exp21
e21g <- readRDS('exp21/exp_21_global.rds')|>
  msns_to_long()
e21p <- readRDS('exp21/exp_21_phospho.rds')|>
  msns_to_long(dtype='phospho')

e21 <- rbind(e21g,e21p)
e21$experiment = 'Patient samples'
e21$Sample <- e21$PatientID

e21$Treatment[e21$Treatment%in%c('BASELINE','BASELINE (No ASO)')] <- 'Control'



#exp22
e22g <- readRDS('exp22/exp_22_global.rds')|>
  msns_to_long()
e22p <- readRDS('exp22/exp_22_phospho.rds')|>
  msns_to_long(dtype='phospho')
e22 <- rbind(e22g,e22p)
e22$Treatment <- rep("Control", nrow(e22))
e22$Treatment[grep('NRAS',e22$`Sample Name`)] <-'NRAS KO'
e22$Treatment[grep('NT guide',e22$`Sample Name`)] <-'NT Guide'
e22$experiment = 'Gilteritinib resistant cell line'

vars <-c('experiment','Treatment','feature', 'Sample', 'data_type','expr')
full_dat<- rbind(select(e20,vars),select(e21,vars), select(e22,vars))

```

Now we can try to plot the plots elie wants

```{r}

mf <- sapply(mono,function(x) grep(paste0('^',x),full_dat$feature))

mono_g <- full_dat[unlist(mf),]

ggplot(mono_g,aes(x=expr, y = feature, fill=Treatment, alpha =0.5)) + ggridges::geom_density_ridges() + facet_grid(~experiment)

ggsave("mono_markers_density.pdf", width=12)



pf <- sapply(prim,function(x) grep(paste0('^',x),full_dat$feature))

#p_g <- full_dat[unlist(pf),]

#ggplot(p_g,aes(x=expr, y = feature, fill=Treatment, alpha =0.5)) + ggridges::geom_density_ridges() + facet_grid(~experiment)

#ggsave("prim_markers_density.pdf")



inf <- sapply(int,function(x) grep(paste0('^',x),full_dat$feature))

i_g <- full_dat[unlist(inf),]

ggplot(i_g,aes(x=expr, y = feature, fill=Treatment, alpha =0.5)) + ggridges::geom_density_ridges() + facet_grid(~experiment)

ggsave("interesting_markers_density.pdf", width=12)


```