---
title: "Dataset Comparisons, Exp24 + Exp27 combined"
author: "Jim Sanford"
date: "2026-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MSnSet.utils)
library(tidyverse)
library(eulerr)
library(ggpubr)
library(readr)
library(pheatmap)
library(synapser)
```

## Downstream analysis and comparison across datasets

Having cleaned up the different versions of the Exp24 + Exp27 combined data in the previous notebook, we will now run some statistics within each dataset to see how they compare to each other. Additionally, within each dataset, we will evaluate how filtering proteins and/or samples affects the resulting statistics.

This first section loads in the median-centered crosstabs generated in the previous script and sample metadata, all of which are stored on synapse. Data and metadata are assembled into msnsets for analysis.

```{r import data, build msnsets for testing, message=FALSE}
#For pulling files from synapse
synapser::synLogin()

#Import a detailed metadata file (generated from Belinda's old code, stored on synapse as syn74023681)
meta <- synapser::synGet("syn74023681")
metadata <- as.data.frame(read_tsv(meta$path))

#Change metadata$patient from numeric to character (so it's not a continuous variable when used as covariate in limma)
metadata$patient <- as.character(metadata$patient)

#make cellsort factor (combining cell type and sort type) for limma testing later
metadata$cellsort <- paste0(metadata$cellType, "_", metadata$Sort.Type)
rownames(metadata) <- metadata$sample

#Build msnsets: first with all proteins (no missingness filter)
#Importing expression-level data generated in previous notebook and now stored on synapse
v1_all <- synapser::synGet("syn74022739")
v1_allprot <- as.matrix(read.delim(v1_all$path))
colnames(v1_allprot) <- gsub("X", "", colnames(v1_allprot))
v1_allprot <- MSnSet(exprs = v1_allprot, pData = metadata)

v2_all <- synapser::synGet("syn74022812")
v2_allprot <- as.matrix(read.delim(v2_all$path))
colnames(v2_allprot) <- gsub("X", "", colnames(v2_allprot))
v2_allprot <- MSnSet(exprs = v2_allprot, pData = metadata)

v3_all <- synapser::synGet("syn74022813")
v3_allprot <- as.matrix(read.delim(v3_all$path))
colnames(v3_allprot) <- gsub("X", "", colnames(v3_allprot))
v3_allprot <- MSnSet(exprs = v3_allprot, pData = metadata)

v4_all <- synapser::synGet("syn74022811")
v4_allprot <- as.matrix(read.delim(v4_all$path))
colnames(v4_allprot) <- gsub("X", "", colnames(v4_allprot))
v4_allprot <- MSnSet(exprs = v4_allprot, pData = metadata)

#Build msnsets: now with filtered proteins (50% missingness filter)
#Importing expression-level data generated in previous notebook and now stored on synapse
v1_filt <- synapser::synGet("syn74022738")
v1_filtprot <- as.matrix(read.delim(v1_filt$path))
colnames(v1_filtprot) <- gsub("X", "", colnames(v1_filtprot))
v1_filtprot <- MSnSet(exprs = v1_filtprot, pData = metadata)

v2_filt <- synapser::synGet("syn74022814")
v2_filtprot <- as.matrix(read.delim(v2_filt$path))
colnames(v2_filtprot) <- gsub("X", "", colnames(v2_filtprot))
v2_filtprot <- MSnSet(exprs = v2_filtprot, pData = metadata)

v3_filt <- synapser::synGet("syn74022810")
v3_filtprot <- as.matrix(read.delim(v3_filt$path))
colnames(v3_filtprot) <- gsub("X", "", colnames(v3_filtprot))
v3_filtprot <- MSnSet(exprs = v3_filtprot, pData = metadata)

v4_filt <- synapser::synGet("syn74022809")
v4_filtprot <- as.matrix(read.delim(v4_filt$path))
colnames(v4_filtprot) <- gsub("X", "", colnames(v4_filtprot))
v4_filtprot <- MSnSet(exprs = v4_filtprot, pData = metadata)
```

Crosstabs generated in the previous script were not filtered for sample-level missingness as was done in prior work. Missingness is evaluated within each version of the data, and an additional msnset is generated by removing samples with <75% coverage. This was done previously, so we want to replicate that for comparison across datasets and with prior results. 75% coverage seems like a pretty conservative filter, might try dropping that to 50% as well.

```{r check out missingness within each sample, filter as done previously, message=FALSE}
#Calculate missingness within samples, filter out  subjects with <75% coverage
x1 <- as.data.frame(colMeans(!is.na(exprs(v1_filtprot))))
x1 <- subset(x1, x1$`colMeans(!is.na(exprs(v1_filtprot)))` <0.75)

v1_filtprot_filtsamp <- v1_filtprot[,!colnames(exprs(v1_filtprot)) %in% rownames(x1)]

x2 <- as.data.frame(colMeans(!is.na(exprs(v2_filtprot))))
x2 <- subset(x2, x2$`colMeans(!is.na(exprs(v2_filtprot)))` <0.75)

v2_filtprot_filtsamp <- v2_filtprot[,!colnames(exprs(v2_filtprot)) %in% rownames(x2)]

x3 <- as.data.frame(colMeans(!is.na(exprs(v3_filtprot))))
x3 <- subset(x3, x3$`colMeans(!is.na(exprs(v3_filtprot)))` <0.75)

v3_filtprot_filtsamp <- v3_filtprot[,!colnames(exprs(v3_filtprot)) %in% rownames(x3)]

x4 <- as.data.frame(colMeans(!is.na(exprs(v4_filtprot))))
x4 <- subset(x4, x4$`colMeans(!is.na(exprs(v4_filtprot)))` <0.75)

v4_filtprot_filtsamp <- v4_filtprot[,!colnames(exprs(v4_filtprot)) %in% rownames(x4)]
```

Previous results are imported from synapse for later comparison with results from these data. Goal is to see how different data processing affects downstream statistical results.

```{r Importing previous stats for downstream comparisons, message=FALSE}
#import Belinda's previous DA results for comparisons; stored on synapse as syn74024893
res_bel <- readxl::read_xlsx(synapser::synGet("syn74024893")$path, 
                                 sheet = "Differential_expression")
res_bel <- filter(res_bel, Contrast == "CD14_Pos_vs_Neg")
colnames(res_bel)[colnames(res_bel) == 'Gene'] <- 'feature'
colnames(res_bel)[colnames(res_bel) == 'Log2FC'] <- 'logFC'
```

Initial limma analysis is run to generate statistics for CD14 vs CD34 cells within each dataset. This is performed on the datasets that were filtered for proteins with <50% missingness, in the following iterations:

  -CD14 vs CD34 comparisons  made with all samples (bead-sorted and flow-sorted, with sort type included as a covariate) or only within bead-sorted samples (to match and compare with previous analysis)
  -with or without removing samples with <75% coverage (msnsets ending in _filtprot or _filtprot_filtsamp)
  
Thus, a total of 4 sets of limma results are generated for each dataset. These will be compared:
  -across datasets (i.e. comparing the results of the same statistical model run on all 4 versions of the data), and
  -within datasets (i.e. comparing the results of the 4 statistical models run within a single version of the data)
  
Ultimately, the results of these comparisons will be used to evaluate which dataset and which model(s) should be used moving forward.

```{r limma statistics, message=FALSE}
#Run some limma: CD14 vs CD34, with or without filtering low-coverage samples (as was done in prior results), in 2 formats:
#1. with all samples included, using sort type (bead or flow) as a covariate
#2. only testing the bead-sorted samples (as was done in prior results)

#First testing all CD14 vs CD34, controlling for sort type in model
contrasts <- c("cellTypeCD14 - cellTypeCD34")

res_v1_allsamples <- limma_contrasts(v1_filtprot,
                                      model.str = "~ 0 + cellType + Sort.Type + patient", 
                                      coef.str = "cellType", 
                                      contrasts = contrasts)

res_v1_filtsamples <- limma_contrasts(v1_filtprot_filtsamp,
                                      model.str = "~ 0 + cellType + Sort.Type + patient", 
                                      coef.str = "cellType", 
                                      contrasts = contrasts)

res_v2_allsamples <- limma_contrasts(v2_filtprot,
                                      model.str = "~ 0 + cellType + Sort.Type + patient", 
                                      coef.str = "cellType", 
                                      contrasts = contrasts)

res_v2_filtsamples <- limma_contrasts(v2_filtprot_filtsamp,
                                      model.str = "~ 0 + cellType + Sort.Type + patient", 
                                      coef.str = "cellType", 
                                      contrasts = contrasts)

res_v3_allsamples <- limma_contrasts(v3_filtprot,
                                      model.str = "~ 0 + cellType + Sort.Type + patient", 
                                      coef.str = "cellType", 
                                      contrasts = contrasts)

res_v3_filtsamples <- limma_contrasts(v3_filtprot_filtsamp,
                                      model.str = "~ 0 + cellType + Sort.Type + patient", 
                                      coef.str = "cellType", 
                                      contrasts = contrasts)

res_v4_allsamples <- limma_contrasts(v4_filtprot,
                                      model.str = "~ 0 + cellType + Sort.Type + patient", 
                                      coef.str = "cellType", 
                                      contrasts = contrasts)

res_v4_filtsamples <- limma_contrasts(v4_filtprot_filtsamp,
                                      model.str = "~ 0 + cellType + Sort.Type + patient", 
                                      coef.str = "cellType", 
                                      contrasts = contrasts)


#Now testing for just bead-sorted CD14 vs CD34 
contrasts2 <- c("cellsortCD14_Bead - cellsortCD34_Bead")

res_v1_allsamples_cellsort <- limma_contrasts(v1_filtprot,
                                               model.str = "~ 0 + cellsort + patient", 
                                               coef.str = "cellsort", 
                                               contrasts = contrasts2)

res_v1_filtsamples_cellsort <- limma_contrasts(v1_filtprot_filtsamp,
                                               model.str = "~ 0 + cellsort + patient", 
                                               coef.str = "cellsort", 
                                               contrasts = contrasts2)

res_v2_allsamples_cellsort <- limma_contrasts(v2_filtprot,
                                               model.str = "~ 0 + cellsort + patient", 
                                               coef.str = "cellsort", 
                                               contrasts = contrasts2)

res_v2_filtsamples_cellsort <- limma_contrasts(v2_filtprot_filtsamp,
                                               model.str = "~ 0 + cellsort + patient", 
                                               coef.str = "cellsort", 
                                               contrasts = contrasts2)

res_v3_allsamples_cellsort <- limma_contrasts(v3_filtprot,
                                               model.str = "~ 0 + cellsort + patient", 
                                               coef.str = "cellsort", 
                                               contrasts = contrasts2)

res_v3_filtsamples_cellsort <- limma_contrasts(v3_filtprot_filtsamp,
                                               model.str = "~ 0 + cellsort + patient", 
                                               coef.str = "cellsort", 
                                               contrasts = contrasts2)

res_v4_allsamples_cellsort <- limma_contrasts(v4_filtprot,
                                               model.str = "~ 0 + cellsort + patient", 
                                               coef.str = "cellsort", 
                                               contrasts = contrasts2)

res_v4_filtsamples_cellsort <- limma_contrasts(v4_filtprot_filtsamp,
                                               model.str = "~ 0 + cellsort + patient", 
                                               coef.str = "cellsort", 
                                               contrasts = contrasts2)
```

To get a sense of how similar the results across datasets look, we calculate correlation of logFC values or t-statistics from each dataset and generate heatmaps. For this, we will just use one of the versions of statistics generated above (no sample filtering based on protein coverage, using all samples (and controlling for sort type) when comparing CD14+ vs CD34+ cell populations.

```{r Heatmaps: comparisons across datasets, message=FALSE}
##Group results together in a list: 4 different datasets all tested the same way
dfs <- list(
  v1 = res_v1_allsamples, 
  v2 = res_v2_allsamples, 
  v3 = res_v3_allsamples,
  v4 = res_v4_allsamples)

#Gather data (logFC or t-stat) and generate correlation matrices
logfc_list <- lapply(dfs, function(x) {
  v <- x$logFC
  names(v) <- x$feature   # adjust if your gene column has another name
  v
})

tstat_list <- lapply(dfs, function(x) {
  v <- x$t
  names(v) <- x$feature   # adjust if your gene column has another name
  v
})

common_genes <- Reduce(intersect, lapply(logfc_list, names))

logfc_mat <- sapply(logfc_list, function(v) v[common_genes])

tstat_mat <- sapply(tstat_list, function(v) v[common_genes])

cor_mat_logFC <- cor(logfc_mat, method = "pearson", use = "pairwise.complete.obs")

cor_mat_tstat <- cor(tstat_mat, method = "pearson", use = "pairwise.complete.obs")

#Generate Heatmaps
breaks <- seq(0.8, 1, length.out = 100)

pheatmap(
  cor_mat_tstat,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  display_numbers = TRUE,   # show r values on the heatmap
  number_format = "%.3f",   # two decimal places
  fontsize_number = 12,
  # number_color = "black",
  color = colorRampPalette(c("blue", "white", "red"))(length(breaks) -1),
  breaks = breaks,
  main = "Correlation Across Datasets, t-statistic",
)

pheatmap(
  cor_mat_logFC,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  display_numbers = TRUE,   # show r values on the heatmap
  number_format = "%.3f",   # two decimal places
  fontsize_number = 12,
  # number_color = "black",
  color = colorRampPalette(c("blue", "white", "red"))(length(breaks) -1),
  breaks = breaks,
  main = "Correlation Across Datasets, log(FC)",
)

```
This comparison shows that the results from the first 3 versions of the data (all generated thru DIA-NN, with differences related to FAIMS CV correction and quantUMS application) correlate fairly well (pearson r >0.91 in all cases); version 4 of the data (generated thru Spectronaut) has lower correlation but not that bad (pearson v >0.84), and this is expected as the data was processed with a different software.

Now, heatmaps are generated for correlation of logFC or t-statistics across the 4 sets of statistics within a particular dataset (using results from "version 2" of the data as a benchmark), to see how removing samples (based on missingness) and comparing all samples vs. only bead-sorted samples affect the results.

```{r Heatmaps: comparisons within datasets, message=FALSE}
##Group results together in a list: 4 different analyses within the same dataset
dfs <- list(
  "All+celltype" = res_v2_allsamples, 
  "Filt+celltype" = res_v2_filtsamples, 
  "All+cellsort" = res_v2_allsamples_cellsort,
  "Filt+cellsort" = res_v2_filtsamples_cellsort)

#Gather data (logFC or t-stat) and generate correlation matrices
logfc_list <- lapply(dfs, function(x) {
  v <- x$logFC
  names(v) <- x$feature   # adjust if your gene column has another name
  v
})

tstat_list <- lapply(dfs, function(x) {
  v <- x$t
  names(v) <- x$feature   # adjust if your gene column has another name
  v
})

common_genes <- Reduce(intersect, lapply(logfc_list, names))

logfc_mat <- sapply(logfc_list, function(v) v[common_genes])

tstat_mat <- sapply(tstat_list, function(v) v[common_genes])

cor_mat_logFC <- cor(logfc_mat, method = "pearson", use = "pairwise.complete.obs")

cor_mat_tstat <- cor(tstat_mat, method = "pearson", use = "pairwise.complete.obs")

#Generate Heatmaps
breaks <- seq(0.9, 1, length.out = 100)

pheatmap(
  cor_mat_tstat,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  display_numbers = TRUE,   # show r values on the heatmap
  number_format = "%.3f",   # two decimal places
  fontsize_number = 12,
  # number_color = "black",
  color = colorRampPalette(c("blue", "white", "red"))(length(breaks) -1),
  breaks = breaks,
  main = "Correlation Within Dataset_v2: t-statistic",
)

pheatmap(
  cor_mat_logFC,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  display_numbers = TRUE,   # show r values on the heatmap
  number_format = "%.3f",   # two decimal places
  fontsize_number = 12,
  # number_color = "black",
  color = colorRampPalette(c("blue", "white", "red"))(length(breaks) -1),
  breaks = breaks,
  main = "Correlation Within Dataset_v2: log(FC)",
)

```
This comparison shows that all 4 statistical models within a dataset produce highly similar results. My preference would be to include as many samples as possible if it's not negatively affecting the results, thus leaning towards using all samples (not filtering on protein coverage) and comparing all CD14+ vs CD34+ populations regardless of sort type (while using sort type as a covariate in linar model).

Now let's make the same comparison and heatmaps as done 2 sections ago (comparison across the 4 different datasets), but now including previous results generated from this data for additional comparison. Previous analysis was performed on data that filtered out samples with <75% protein coverage, and only used bead-sorted samples for CD14+ vs CD34+ comparisons, so we will use that version of our statistical analysis for this comparison (results sets ending in _filtsamples_cellsort).

```{r Heatmaps: comparisons across datasets + older results, message=FALSE}
##Group results together in a list: 4 different datasets all tested the same way
dfs <- list(
  v1 = res_v1_filtsamples_cellsort, 
  v2 = res_v2_filtsamples_cellsort, 
  v3 = res_v3_filtsamples_cellsort,
  v4 = res_v4_filtsamples_cellsort,
  old = res_bel)

#Gather data (logFC or t-stat) and generate correlation matrices
logfc_list <- lapply(dfs, function(x) {
  v <- x$logFC
  names(v) <- x$feature   # adjust if your gene column has another name
  v
})

tstat_list <- lapply(dfs, function(x) {
  v <- x$t
  names(v) <- x$feature   # adjust if your gene column has another name
  v
})

common_genes <- Reduce(intersect, lapply(logfc_list, names))

logfc_mat <- sapply(logfc_list, function(v) v[common_genes])

tstat_mat <- sapply(tstat_list, function(v) v[common_genes])

cor_mat_logFC <- cor(logfc_mat, method = "pearson", use = "pairwise.complete.obs")

cor_mat_tstat <- cor(tstat_mat, method = "pearson", use = "pairwise.complete.obs")

#Generate Heatmaps
breaks <- seq(0.8, 1, length.out = 100)

pheatmap(
  cor_mat_tstat,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  display_numbers = TRUE,   # show r values on the heatmap
  number_format = "%.3f",   # two decimal places
  fontsize_number = 12,
  # number_color = "black",
  color = colorRampPalette(c("blue", "white", "red"))(length(breaks) -1),
  breaks = breaks,
  main = "Correlation Across Datasets, t-statistic",
)

pheatmap(
  cor_mat_logFC,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  display_numbers = TRUE,   # show r values on the heatmap
  number_format = "%.3f",   # two decimal places
  fontsize_number = 12,
  # number_color = "black",
  color = colorRampPalette(c("blue", "white", "red"))(length(breaks) -1),
  breaks = breaks,
  main = "Correlation Across Datasets, log(FC)",
)

```
Not awful, and probably expected, but the previous results show the weakest correlations across these 5 datasets. Data was processed differently previously (namely that samples from Exp24 and Exp27 were analyzed in separate batches in DIA-NN), so this makes sense, and the pearson r values are still relatively high (>0.8 in all comparisons).

As an additional level of comparison of results across the various datasets, we can generate UpSet plots showing the overlap of a) all proteins identified/ tested in each dataset, and b) statistically significant (adjusted p-value <0.05) proteins from CD14 vs CD34 comparisons within each dataset.

```{r UpSet Plots: comparisons across datasets, message=FALSE}
##Group results together in a list: 4 different datasets all tested the same way
dfs <- list(
  v1 = res_v1_allsamples, 
  v2 = res_v2_allsamples, 
  v3 = res_v3_allsamples,
  v4 = res_v4_allsamples)

#Extract lists of genes from each dataset (all genes tested; all significant genes)
gene_lists <- lapply(dfs, function(x) {
  x$feature   # change padj / gene if needed
})

sig_lists <- lapply(dfs, function(x) {
  subset(x, adj.P.Val < 0.05)$feature   # change padj / gene if needed
})

# Convert list of gene sets to an UpSetR-compatible format
upset_input_sig <- UpSetR::fromList(sig_lists)

upset_input_all <- UpSetR::fromList(gene_lists)

# Create UpSet plot
desired_order <- c("v4", "v3", "v2", "v1")

UpSetR::upset(
  upset_input_sig,
  sets = desired_order,
  keep.order = T,
  nsets = length(sig_lists),      # number of sets to show
  # nintersects = 10,               # show all intersections (can limit if too many)
  order.by = "freq",              # order intersections by size
  sets.x.label = "Significant Genes (p.adj<0.05)",
  mainbar.y.label = "Intersection size",
  # set_size.show = TRUE,
  text.scale = c(1.5, 1.5, 1, 1.25, 1.5, 1.5)  # adjust text size if needed
)

UpSetR::upset(
  upset_input_all,
  sets = desired_order,
  keep.order = T,
  nsets = length(sig_lists),      # number of sets to show
  # nintersects = 10,               # show all intersections (can limit if too many)
  order.by = "freq",              # order intersections by size
  sets.x.label = "Identified Proteins",
  mainbar.y.label = "Intersection size",
  # set_size.show = TRUE,
  text.scale = c(1.5, 1.5, 1, 1.25, 1.5, 1.5)  # adjust text size if needed
)
```
The overall consistency of the DEGs from the different datasets here is promising. As seen in the earlier correlations, versions 1-3 look the most similar. Each of the 4 datasets has a set of ~200-300 "unique" DEGs (some of these could be proteins that were uniquely identified in any given dataset), with version 4 of the data having the largest unique set. Version 2 of the data yields the highest number of DEGs from this statistical model, perhaps indicating the the FAIMS CV correction employed by James Fulcher in this version reduces variability in the data.


Again, generating UpSet plots to compare results, this time using the 4 different sets of limma results generated from a single dataset (using results from "version 2" of the data as a benchmark).

```{r UpSet Plots: comparisons within datasets, message=FALSE}
##Group results together in a list: 4 different analyses within the same dataset
dfs <- list(
  "All+celltype" = res_v2_allsamples, 
  "Filt+celltype" = res_v2_filtsamples, 
  "All+cellsort" = res_v2_allsamples_cellsort,
  "Filt+cellsort" = res_v2_filtsamples_cellsort)

#Extract lists of significant genes from each dataset
sig_lists_v2 <- lapply(dfs, function(x) {
  subset(x, adj.P.Val < 0.05)$feature   # change padj / gene if needed
})

# Convert list of gene sets to an UpSetR-compatible format
upset_input_v2 <- UpSetR::fromList(sig_lists_v2)

# Create UpSet plot
desired_order <- c("Filt+cellsort", "All+cellsort", "Filt+celltype", "All+celltype")

UpSetR::upset(
  upset_input_v2,
  sets = desired_order,
  keep.order = T,
  nsets = length(sig_lists),      # number of sets to show
  # nintersects = 10,               # show all intersections (can limit if too many)
  order.by = "freq",              # order intersections by size
  sets.x.label = "Significant Genes (p.adj<0.05)",
  mainbar.y.label = "Intersection size",
  # set_size.show = TRUE,
  text.scale = c(1.5, 1.5, 1, 1.25, 1.5, 1.5)  # adjust text size if needed
)
```
Good overlap of significant genes from the different models within a dataset here, but a few things look slightly suspicious:
  -Using the patient-filtered dataset and all samples (Filt + celltype) gives ~1000 more DEGs than any other model
  -Using all patient samples and only comparing bead-sorted samples (All+cellsort) gives ~1000 DEGs less than others
As we know the logFC and t-statistics from the 4 models correlate well (based on heatmaps from above), I assume this is based on slight differences from p-value adjustment, and a full rank-based analysis of the results (GSEA or CAMERA-PR) would produce highly similar results from all 4 models. However, using a threshold-based approach (ORA) would likely yield different results if ~1000 genes were included or excluded from the input lists.


Out of curiosity, evaluating how limma results are affected if patient ID is not included as a covariate in the linear model. Generate another set of results from version 2 of the data and compare with results generated above.

```{r testing using patient as covariate vs not, message=FALSE}
####Generate one more results set, without using patient as covariate
##Will just use v2 of the data, all samples, comparing cellType
res_v2_allsamples_nocovariate <- limma_contrasts(v2_filtprot,
                                     model.str = "~ 0 + cellType + Sort.Type", 
                                     coef.str = "cellType", 
                                     contrasts = contrasts)

colnames(res_v2_allsamples_nocovariate) <- paste0(colnames(res_v2_allsamples_nocovariate), "_v2")
colnames(res_v2_allsamples_nocovariate)[colnames(res_v2_allsamples_nocovariate) == 'feature_v2'] <- 'feature'

z <- full_join(res_v2_allsamples, res_v2_allsamples_nocovariate)

ggscatter(z, x='logFC', y='logFC_v2', add = 'reg.line',
          add.params = list(color = 'blue', fill = 'lightgray'),
          conf.int = TRUE, cor.coef = TRUE, cor.coeff.args = list(method= 'pearson'))+
  labs(title = "Sorted Cell Proteomics DA Comparisons",
       subtitle = "All samples, testing cell type",
       x = "log(FC), patient covariate",
       y = "log(FC), no patient covariate") +
  geom_vline(xintercept = 0, linetype = 2, color = "red") +
  geom_hline(yintercept = 0, linetype = 2, color = "red") +
  geom_abline(intercept = 0, slope = 1, color = "green", linetype = "dashed")


ggscatter(z, x='t', y='t_v2', add = 'reg.line',
          add.params = list(color = 'blue', fill = 'lightgray'),
          conf.int = TRUE, cor.coef = TRUE, cor.coeff.args = list(method= 'pearson'))+
  labs(title = "Sorted Cell Proteomics DA Comparisons",
       subtitle = "All samples, testing cell type",
       x = "t-stat, patient covariate",
       y = "t-stat, no patient covariate") +
  geom_vline(xintercept = 0, linetype = 2, color = "red") +
  geom_hline(yintercept = 0, linetype = 2, color = "red") +
  geom_abline(intercept = 0, slope = 1, color = "green", linetype = "dashed")

```
Very little differences in the results with or without using patient as a covariate, though it does seem like not including patient results in smaller calculated logFC and t-statistic (as judged by the axis scales and the dashed green line with a slope of 1). I intend to include patient in the models moving forward.

Saving some relevant objects (msnsets and limma results) for use in the next script.

```{r saving data for use in next notebook, message=FALSE}
save(v2_allprot,
     v2_filtprot,
     v2_filtprot_filtsamp,
     res_v2_allsamples,
     res_v2_allsamples_cellsort,
     res_v2_filtsamples,
     res_v2_filtsamples_cellsort,
     file = "v2_msnsets_stats.RData")
```