---
title: "myelo_nglycan_analysis"
output: html_document
---

Here we evaluate the N-glycan patterns from the patient samples.

```{r load, warning=FALSE, message=FALSE, echo = FALSE}
library(tidyverse)
library(synapser)
library(spammR)
library(ggplot2)
synLogin()
```

# Load data into SpatialExperiment object

First we get the data into the correct format
```{r load data}

#data from brittney
load(synapser::synGet('syn77540088')$path)
#metadta
load(synapser::synGet('syn77540060')$path)

metadata <- t(metadata) |>
  as.data.frame() |>
  mutate(x = as.numeric(x),
         y = as.numeric(y),
         z = as.numeric(z))

#n-glycan annotation from Geremy's code
source("https://raw.githubusercontent.com/GeremyClair/NglycanMiniOn/refs/heads/main/R/01_base_functions.R")

##find this
fmet <- data.frame(nGlycan = rownames(data))

ng <- NGlycan_miner(rownames(data)) |>
  dplyr::rename(nGlycan='ID')|>
  right_join(fmet)

#join with data and add to object
rownames(ng) <- rownames(data)
spe <- spammR::convert_to_spe(dat=data, 
                              sample_meta = metadata,
                              spatial_coords_colnames = c("x","y"),
                              feature_meta =ng)


```

# Calculate missingness across samples and N-glycans

How many samples/points/glycans have a high degree of missingness?

```{r calculate misingness}

##per feature missiningness
missing <- calc_missingness(spe)

##calculate per sample missingness
missing <- calc_missingness(missing, group_name = 'Donor_Time')


md <- as.data.frame(rowData(missing)) |> ##now we can try to plot missingness by
  tidyr::pivot_longer(starts_with('missingSample'),names_to='condition',
                      values_to='fracMissing') |>
  tidyr::separate(col='condition',sep='_',into=c('missing','sample','condition'))

md|>
  subset(is.na(condition))|>
  ggplot(aes(x=fracMissing, fill= Tetraantenary)) + geom_histogram()

md |>
  subset(!is.na(condition)) |>
  ggplot(aes(x=fracMissing,fill=sample)) + 
  geom_histogram(position='dodge') +
  facet_grid(condition~.) 

```
There are more missing N-glycans post transplant. However, more are 100% missing.
What if we remove those that are missing across an entire sample? 

```{r missing issues}

md|>subset(fracMissing == 1.0) |> 
  group_by(sample,condition)|> 
  summarize(nMissing = n()) |>
  ggplot(aes(x=sample, fill=condition, y= nMissing)) +
  geom_bar(position='dodge', stat='identity')

```

Weirdly not all n-glycans are missing post-treatment, some are missing
pre-treatment. Patient 3 has no pre-transplant sample. 

Which glycan are missing in both conditions, or one?

```{r subset and plot missingness, warning = FALSE}

missing <- md |> subset(fracMissing == 1.0) |> 
  dplyr::select('nGlycan','sample','condition','fracMissing') |> 
  tidyr::pivot_wider(names_from='condition',values_from='fracMissing') |>
  mutate(missingCondition = ifelse(!is.na(post.transplant),  #checking to see if there is 100% missingness. if NA it means its not missing. 
                                    ifelse(!is.na(pre.transplant), 
                                           'Missing in both','Missing Post'), 
                                   ifelse(!is.na(pre.transplant),'MissingPre','Present in both'))) |>
  dplyr::select(nGlycan,sample, missingCondition) |>
  right_join(md) |>
  replace_na(replace = list(missingCondition='Present in both'))

missing |>
  subset(!is.na(sample)) |># remove bulk
ggplot(aes(x=missingCondition,fill=sample)) + 
  geom_histogram(position='dodge',stat='count')
```
Let's figure out how many of these are shared across samples.

```{r missing counts}

similar <- missing |> subset(!is.na(sample))|>
  #subset(sample!=3) |>
    group_by(nGlycan, missingCondition) |> 
    summarize(numSamples = n()) |>
    arrange(desc(numSamples)) |>
    subset(numSamples %in% c(6,11))

ggplot(similar, aes(x=missingCondition)) + geom_histogram(stat = 'count')
```
What are the glycans that are missing post-treatment systematically?

```{r}

subset(similar, missingCondition == 'Missing Post')


```

Not really! only these four!

## Abundance changes pre-post treatment

Let's do standard differential expression pre/post treatment. We'll reduce
to only those N-glycans that are present in both conditions in at least 
one sample (otherwise the fold changes end up being Infinite)

```{r look at differential expression, enrichment, warning=FALSE}

ngs <- missing |>
  subset(!is.na(sample)) |>
  subset(sample!=3) |>
  subset(missingCondition == 'Present in both') 

print(paste("We found",length(unique(ngs$nGlycan)),"nGlycans that are present in both pre- and post-treated samples"))

rspe <- spe[rowData(spe)[,'nGlycan']%in%ngs$nGlycan,]
res <- calc_spatial_diff_ex(rspe, count_based = FALSE,
                            log_transformed = TRUE, category_col = 'Time')

#subset(rowData(res), `pre-transplant_vs_post-transplant.adj.P.Val.limma` < 0.05)

volcano_plot(res, logFC_colname ='pre-transplant_vs_post-transplant.logFC.limma',
             pval_colname = 'pre-transplant_vs_post-transplant.adj.P.Val.limma',
             sigLabel_colname = 'nGlycan',
             pval_corrected = TRUE,
             log_transformed = TRUE,
             thresh = 0.05,
             title='Pre vs post transplant, no imputation')

```
This looks super suspicious. Let's try some imputation perhaps. what if we only
use the most expressed Nglycans?

```{r restrict}

present <- similar|> subset(missingCondition=='Present in both')


rspe <- spe[rowData(spe)[,'nGlycan']%in%present$nGlycan,]
res <- calc_spatial_diff_ex(rspe, count_based = FALSE,
                            log_transformed = FALSE, category_col = 'Time')

#subset(rowData(res), `pre-transplant_vs_post-transplant.adj.P.Val.limma` < 0.05)

volcano_plot(res, logFC_colname ='pre-transplant_vs_post-transplant.logFC.limma',
             pval_colname = 'pre-transplant_vs_post-transplant.adj.P.Val.limma',
             sigLabel_colname = 'nGlycan',
             pval_corrected = TRUE,
             log_transformed = FALSE,
             thresh = 0.05,
             title='Pre vs post transplant, present only')

```

### How does imputation change this?

```{r 0 imputation}

res <- impute_spe(spe, method = "zero")
res <- calc_spatial_diff_ex(res, count_based = FALSE,
                            assay_name = 'imputed',
                            log_transformed = TRUE, category_col = 'Time')

volcano_plot(res, logFC_colname ='pre-transplant_vs_post-transplant.logFC.limma',
             pval_colname = 'pre-transplant_vs_post-transplant.adj.P.Val.limma',
             sigLabel_colname = 'nGlycan',
             pval_corrected = TRUE,
             log_transformed = TRUE,
             thresh = 0.05,
             title='Pre vs post transplant, zero imputation')
```
Same mess of inflated p-values


```{r median imputation}
res <- impute_spe(spe, method = "median")
res <- calc_spatial_diff_ex(res, count_based = FALSE,
                            assay_name = 'imputed',
                            log_transformed = TRUE, category_col = 'Time')

volcano_plot(res, logFC_colname ='pre-transplant_vs_post-transplant.logFC.limma',
             pval_colname = 'pre-transplant_vs_post-transplant.adj.P.Val.limma',
             sigLabel_colname = 'nGlycan',
             pval_corrected = TRUE,
             log_transformed = TRUE,
             thresh = 0.05,
             title='Pre vs post transplant, median imputation')
```



```{r neighbor imputation}
res <- impute_spe(spe, method = "spatial_knn", k = 8)
res <- calc_spatial_diff_ex(res, count_based = FALSE,
                            assay_name = 'imputed',
                            log_transformed = TRUE, category_col = 'Time')

volcano_plot(res, logFC_colname ='pre-transplant_vs_post-transplant.logFC.limma',
             pval_colname = 'pre-transplant_vs_post-transplant.adj.P.Val.limma',
             sigLabel_colname = 'nGlycan',
             pval_corrected = TRUE,
             log_transformed = TRUE,
             thresh = 0.05,
             title='Pre vs post transplant, spatial imputation')
```

## Per patient sample analysis

The overall analysis is too wonky, so let's subset per patient (aside from #3) and
see if we find any trends?

```{r per patient diffex, warning = FALSE}

plots <- lapply(c('1','2','4','5','6'), function (d){
  print(d)
  pspe <- calc_spatial_diff_ex(spe[,spe$Donor == d], count_based =FALSE,
                              compare_vals = c('pre-transplant','post-transplant'),
                               category_col = 'Time', log_transformed = TRUE)
  
  if(length(grep('^pre-',colnames(rowData(pspe)))) >0)
    comp = 'pre-transplant_vs_post-transplant'
  else
    comp = 'post-transplant_vs_pre-transplant'
  volcano_plot(pspe, 
               logFC_colname =paste0(comp,'.logFC.limma'),
             pval_colname = paste0(comp,'.adj.P.Val.limma'),
             
             sigLabel_colname = 'nGlycan',
             pval_corrected = TRUE,
             log_transformed = TRUE,
             thresh = 0.05,
             title=paste('Pre vs post transplant, donor',d))
})

```


## Enrichment

We need to format the glycan dictionary as a gene list, then use leapR to
look at differential enrichment. However, I'm not sure about the data just 
yet.


## Differential spatial activity? 

Again, the data seems to require more processing to do this.
How do we asses differential spatial activity? We

